---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

An introduction of the data and a description of the trends/books/items you are choosing to analyze (and why!)

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

These will likely be calculated using your DPLYR skills, answering questions such as:

- What is the average number of checkouts for each item?
- What is the month or year with the most/least checkouts for a book that you're interested in?
- What is the month or year with the most/least checkouts for ebooks?
- How has the number of print book checkouts changed over time?

Feel free to calculate and report values that you find relevant.

### The Dataset

- Who collected/published the data?
  - The data is published and updated monthly by The Seattle Public Library.

- What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?
  - There are 12 columns in the dataset: the form (digital/physical) of the item, the tool used to checkout, the type of item checked out, year and month of checkout, number of times the title was checked out within a month, and the title, ISBNs, author, subject, publisher, and publication year of the item. For my analysis, I will only use the data between 2017 to 2022 because the data for 2023 is too little to use. I will only use the data that was recorded for the whole year.

- How was the data collected or generated?  
  - The data was collected and generated by The Seattle Public Library as they checkout items.

- Why was the data collected?  
  - The data was collected to not only store the record of checkout but also to provide data
for others to analyze the pattern and trend of item popularity.

- What, if any, ethical questions do you need to consider when working with this data?  
  - Since I don't see any personal information about the people who check out the item, I think the ethical problem for user privacy is not that serious. However, the library user might not be aware that the library is collecting this checkout data. For further analysis, we, or the library, should consider asking for users' consent for the public to see and use the data without ethical issue.

- What are possible limitations or problems with this data?   (at least 200 words)
  - As I look into the dataset, the syntax of the value of each column varies. For example, for the publication year, there are data values like "2022", "[2022]", or "2022,". It will need extra data cleaning in order to process and analyze the data without getting into errors or problems. And for columns like item title, creator, subjects, and publisher, the value is unique since there are so many different items. Unlike checkout year, we can hardly find patterns with these columns. Therefore, we should carefully select the columns or variables that we want to analyze. Additionally, the data has a wide time range, and the data can be affected by some events that happened in these timeframes. For example, some influential events like COVID, new technology released, or events happening in the Seattle area can affect the data. Thus we can analyze or categorize the data based on these events. By doing so, we should also include information about these possible influential factors and explain how the situation would affect the data. Also, the data are only collected in the Seattle Public Library, so it cannot represent its result to the whole Seattle area. It can only tell the situation in its own library. 

### First Trends Over Time Chart

Chart: Line Plot
Question: What is the trend in total checkouts for digital and physical items over the years?
Description: 
This line plot shows the trend of the total checkouts for digital and phtsical items between 2017 to 2022.

``````{r, echo = FALSE, warning = FALSE, code = readLines("chart1.R")}

```

### Second Trends Over Time Chart

Chart: Line Plot
Question: How has the number of average checkouts for the top 3 material types changed over time in 2022?
Description: 

``````{r, echo = FALSE, warning = FALSE, code = readLines("chart2.R")}

```

### Your Choice

Chart: Bar Plot
Question: On average, which year has the highest checkout numbers? How has the main checkout type changed?
Description: 

``````{r, echo = FALSE, warning = FALSE, code = readLines("chart3.R")}

```


